Diffantom: a software stack to build digital phantoms of whole-brain diffusion MRI

### 2015-08-12 paper/main.tex
Remove software layer concept. Close #13
hindered vs restricted diffusion. Close #18

\title[Diffantom]{Diffantom: a software layer to build digital phantoms of whole-brain diffusion MRI}

@@

\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
The simulation process relies on a microstructural model derived from real data (\autoref{fig:figure1}).
On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the
  free-diffusion compartments.
These fraction maps are obtained from the macroscopic structure of tissues in the brain,
  in the following order~%
\footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
  of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
\gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
\footnote{Since here we simulate healthy subjects, the last fraction map is empty and can be omitted}.
On the other hand, the hindered-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
  of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.

@@

  \begin{align*}
  T_1 &= (1.0-f_{cgm}) \cdot T'_1 \\
  T_2 &= (1.0-f_{dgm}) \cdot T'_2 \\
  T_3 &= (1.0-f_{wm}) \cdot T'_3 \\
  T_4 &= T'_4 \\
  T_5 &= 0.0
  \end{align*}
where $f_{\{cgm, dgm, wm\}}$ are the fractions of hindered diffusion for each tissue.
\cite{sepehrband_brain_2015} found out that the fiber fraction ranges across the corpus
  callosum from the 70$\pm$8\% in its body to an upper bound of 80$\pm$11\% in the splenium.
Therefore, we choose $f_{wm} =$ 80\% as default fraction of hindered diffusion in the
  \gls*{wm}.
To our knowledge, hindered diffusion fractions have been studied only for \gls*{wm}.
Therefore, we set $f_{cgm} =$ 25\% and $f_{dgm} =$ 50\% as they yield plausible \gls*{fa}
  and \gls*{adc} maps, assessed visually.
The final $\{F_{1\,\cdots\,3}\}$ maps are computed as follows:

@@

\begin{figure}[h!]
\begin{center}
% \includegraphics[width=\linewidth]{figures/figure01A}
% \includegraphics[width=\linewidth]{figures/figure01B}
\includestandalone[width=\linewidth]{figure01}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
{\textbf{Generation of \emph{diffantoms}:} First, a microstructural model is generated as described
  in the \nameref{sec:data_generation} subsection.
Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
This microstructural model is fed into a simulation block that generates the final \gls*{dmri} dataset
  as described in the second block.
}
\end{figure}

>>>>

\title[Diffantom]{Diffantom: a software stack to build digital phantoms of whole-brain diffusion MRI}

@@

\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
The simulation process relies on a microstructural model derived from real data.
On one hand, the software requires up to five fraction maps $\{T_{1\,\cdots\,5}\}$ of
  free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
These compartements will be derived from the macroscopic structure of tissues within the brain,
  specified in the following order~%
\footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
  of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
\gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
\footnote{Since here we simulate healthy subjects, the last fraction map is empty and can be omitted}.
On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
  of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.

@@

  \begin{align*}
  T_1 &= (1.0-f_{cgm}) \cdot T'_1 \\
  T_2 &= (1.0-f_{dgm}) \cdot T'_2 \\
  T_3 &= (1.0-f_{wm}) \cdot T'_3 \\
  T_4 &= T'_4 \\
  T_5 &= 0.0
  \end{align*}
where $f_{\{cgm, dgm, wm\}}$ are the fractions of restricted diffusion for each tissue.
\cite{sepehrband_brain_2015} found out that the fiber fraction ranges across the corpus
  callosum from the 70$\pm$8\% in its body to an upper bound of 80$\pm$11\% in the splenium.
Therefore, we choose $f_{wm} =$ 80\% as default fraction of restricted diffusion in the
  \gls*{wm}.
To our knowledge, restricted diffusion fractions have been studied only for \gls*{wm}.
Therefore, we set $f_{cgm} =$ 25\% and $f_{dgm} =$ 50\% as they yield plausible \gls*{fa}
  and \gls*{adc} maps, assessed visually.
The final $\{F_{1\,\cdots\,3}\}$ maps are computed as follows:

@@

\begin{figure}[h!]
\begin{center}
\includestandalone[width=\linewidth]{figure01}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
{\textbf{Generation of \emph{diffantoms}:} First, a microstructural model is generated as described
  in the \nameref{sec:data_generation} subsection.
Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
This microstructural model is fed into a simulation block that generates the final \gls*{dmri} dataset
  as described in the second block.
}
\end{figure}
